---
title: "Wrangling with Authenticity"
author: "Prof. Josh Clinton"
institute: "Vanderbilt University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

# Learning Objectives

- Identify the paper's causal claim and how its treatment and outcome are
  measured.
- Determine what comparison would identify the claimed effect.
- State the pattern we should observe before inspecting the data.
- Use `count`, `mutate`, `filter`, `group_by`, and `summarize` to investigate
  the posted data.
- Explain how decisions about cleaning, recoding, and variable names affect an
  analysis.

Today we will investigate a published experimental finding. We will begin with
what the authors claim, identify what pattern we should observe if that claim
is correct, and then inspect the posted data.

The main question for today is:

> Do the posted data provide credible evidence that inauthenticity increases
> the desire to clean?

# The Claim

Read “The Moral Virtue of Authenticity: How Inauthenticity Produces Feelings of
Immorality and Impurity” by Business School professors at Harvard,
Northwestern, and Columbia. (Not too shabby!)

The paper makes a striking claim:

> “Inauthenticity makes people feel morally tainted and leads to a greater
> desire for cleanliness.” (p. 993)

The authors describe the contribution this way:

> “Our research provides the first empirical demonstration that there is indeed
> a link between authenticity and morality.” (p. 994)

What does that mean? Put the claim in your own words before looking at the
data.

YOUR ANSWER HERE:

# Study 4: Inauthenticity Is Not Dissonance

The paper reports five experiments. We will focus on Study 4.

You can skim most of the article. I do not expect you to understand every part
of the analysis. Focus on what the researchers did and what we should observe
if their argument is right.

For the full wording of the study protocol, consult the
[Materials and Methods](readings/Materials_and_methods.docx). A focused summary
of the Experiment 4 materials appears in the appendix.

## Whose Opinions Are Being Studied?

Who participated in Study 4? Is this a random sample of the United States?
People approached at a shopping mall? People who opted into an online survey?
Something else?

YOUR ANSWER HERE:

Does the sample matter for evaluating the claim? Why or why not?

YOUR ANSWER HERE:

## What Did the Researchers Ask People to Do?

Participants first reported whether they supported including course-difficulty
ratings in Harvard’s Q Guide. They were then assigned to one of three writing
conditions.

What were the three conditions? What did participants write in each one?

YOUR ANSWER HERE:

Do the authors have a group that did not write an essay at all? Why might that
matter?

YOUR ANSWER HERE:

## What Is the Outcome?

After writing the essay, participants rated the desirability of five cleaning
products: Dove shower soap, Crest toothpaste, Windex cleaner, Tide detergent,
and Lysol disinfectant.

The variable `av_products_clean` is the average of those five ratings.

Why might the authors use interest in cleaning products as an outcome? Do you
find it convincing as a measure of a “desire for cleanliness”?

YOUR ANSWER HERE:

## What Pattern Should We Expect?

Before looking at the data, state the expected pattern.

Who should report the greatest interest in cleaning products if the paper’s
argument is correct? Who should report the least?

YOUR ANSWER HERE:

This is important. If we do not say what pattern we expect before opening the
data, it becomes very easy to tell a story about whatever pattern we happen to
find.

# Investigating the Data

## Load the Library

What library do we need?

Reference:

```r
library(package_name)
```

Copy the pattern and replace `package_name` with the library we have used in
the previous lectures.

```{r load-library}
# INSERT CODE HERE
```

## Load the Posted Data

The teaching file is named
`GKG_Study4_Class.csv`. It contains the authors’ posted
Study 4 data. The only change from the preserved source CSV is removal of
trailing white space from the open-ended `yearSchool` responses.

What function should we use to read a CSV file? Load the file and save it to an
object with a sensible name.

Reference:

```r
example_data = read.csv("example_file.csv")
```

Copy the pattern. Change the object name and filename to match this study.

```{r load-data}
# INSERT CODE HERE
```

How many rows and columns does the data contain? What are the variable names?

Reference:

```r
dim(example_data)
names(example_data)
```

```{r inspect-data}
# INSERT CODE HERE
```

## What Did Respondents Say?

The survey asked:

> Year in School: ________

Count the responses to `yearSchool`.

Reference:

```r
example_data |>
  count(example_variable)
```

```{r count-year-school}
# INSERT CODE HERE
```

Besides the fact that not all Harvard students can spell, notice anything
weird?

What makes the response “Harvard” different from “Freshman,” “Sophomore,”
“Junior,” or “Senior”? Is “Harvard” an impossible answer, or the answer to
the wrong question?

YOUR ANSWER HERE:

### Turning an Observation Into a Variable

Create a variable equal to 1 for the responses `Harvard` or `harvard` and 0 for
all other responses.

What should we call it?

- `HarvardResponse` describes what appears in the data.
- `WrongAnswer` interprets the response as answering the wrong question.
- `Weird` makes a broader judgment about the observation.
- `Suspicious` makes an even stronger judgment.

Variable names are not always neutral. A name can describe what we observed or
embed our interpretation of what it means.

For now, call the variable `Weird` so that it matches the question we are
asking in this workshop. But be clear about the choice we are making.

Reference:

```r
example_data = example_data |>
  mutate(
    NewVariable = if_else(
      old_variable == "First value" | old_variable == "Second value",
      1,
      0
    )
  )
```

What needs to change? The data object, the new variable name, the old variable,
and the two values we want to identify.

```{r create-weird}
# INSERT CODE HERE
```

Confirm that the variable worked. How many observations are coded as weird?

Use the same `count()` pattern from above. What variable should replace
`example_variable`?

```{r check-weird}
# INSERT CODE HERE
```

## Do the Weird Cases Look Different?

What is the average value of `av_products_clean` among the weird cases?

Reference:

```r
example_data |>
  filter(group_variable == 1) |>
  summarize(
    N = n(),
    AvgOutcome = mean(outcome_variable)
  )
```

Copy and adjust the data, group variable, and outcome variable.

```{r weird-mean}
# INSERT CODE HERE
```

Now calculate the same mean for the non-weird cases.

Return to the code you just ran. Only one value needs to change.

```{r nonweird-mean-first-try}
# INSERT CODE HERE
```

What happened?

If R returned `NA`, do not immediately add code until the result disappears.
First determine why the answer is missing.

Reference:

```r
example_data |>
  filter(is.na(outcome_variable)) |>
  select(variable_one, variable_two, outcome_variable)
```

Which variables would help us understand the missing observation?

```{r inspect-missing-outcome}
# INSERT CODE HERE
```

How many values of `av_products_clean` are missing? Is the missing observation
weird or non-weird? Which condition is it in?

YOUR ANSWER HERE:

Now calculate the non-weird mean using `na.rm = TRUE`.

Reference:

```r
example_data |>
  filter(group_variable == 0) |>
  summarize(
    N = n(),
    Missing = sum(is.na(outcome_variable)),
    AvgOutcome = mean(outcome_variable, na.rm = TRUE)
  )
```

```{r nonweird-mean-corrected}
# INSERT CODE HERE
```

What exactly did `na.rm = TRUE` do? Why is it reasonable here? When might
silently removing missing values be dangerous?

YOUR ANSWER HERE:

# Return to the Paper’s Argument

The comparison above ignores the writing condition. But the paper’s argument
is about what happens when people write something contrary to their stated
beliefs.

Count the values of `condition` and return to the three conditions described in
the paper.

Use the `count()` reference from the year-in-school question. Change the
variable being counted.

```{r count-condition}
# INSERT CODE HERE
```

The two counterattitudinal conditions ask participants to argue against their
stated position. The proattitudinal condition asks them to argue for it.

Create a variable equal to 1 for the two counterattitudinal conditions and 0
for the proattitudinal condition.

The original workshop called this variable `Dissonance`. Is that only a label,
or does it interpret what participants experienced? Would
`CounterAttitudinal` make a different claim?

We will retain `Dissonance` to examine the paper on its own terms. Remember
that the name contains an interpretation.

Reference:

```r
example_data = example_data |>
  mutate(
    NewVariable = if_else(old_variable != "Comparison value", 1, 0)
  )
```

Here `!=` means "not equal to." Which condition is the comparison value?

```{r create-dissonance}
# INSERT CODE HERE
```

## Was the Weird Pattern Concentrated in One Condition?

First count weird and non-weird observations by `Dissonance`.

Reference:

```r
example_data |>
  count(first_group, second_group)
```

```{r count-weird-dissonance}
# INSERT CODE HERE
```

Were the weird observations all assigned to the same condition? Does the count
by itself explain the pattern?

YOUR ANSWER HERE:

### What Should We Compare?

We now have two variables:

- `Weird`: whether the year-in-school response was Harvard.
- `Dissonance`: whether the assigned essay was counterattitudinal.

We want to compare the average cleaning-product rating for all four
combinations.

Before running the code, write down the pattern the paper predicts. Then ask a
more specific question: what pattern among the weird observations would make
you concerned?

YOUR ANSWER HERE:

Use `group_by()` and `summarize()` to report the number of observations, the
number of missing outcomes, and the mean cleaning-product rating for each
combination.

Reference:

```r
example_data |>
  group_by(first_group, second_group) |>
  summarize(
    N = n(),
    Missing = sum(is.na(outcome_variable)),
    AvgOutcome = mean(outcome_variable, na.rm = TRUE),
    .groups = "drop"
  )
```

Now do it!

```{r four-cell-comparison}
# INSERT CODE HERE
```

### What Do You Observe?

Do not move on yet. Read the four rows carefully.

- What is the mean among weird respondents in the dissonance condition?
- What is the mean among weird respondents outside the dissonance condition?
- How do those means compare with the possible range of 1 through 7?
- How many observations produce this pattern?
- Is the same pattern present among the non-weird observations?

YOUR ANSWER HERE:

## Does the Pattern Support the Paper Too Well?

The suspicious observations do not merely support the paper’s argument. They
produce an almost perfect version of it.

What would a respondent have to enter on the five cleaning-product questions
to receive an average of exactly 7? What would an average near 2.4 require?

Could this pattern have occurred by chance? Yes. But is chance a convincing
explanation for what we see?

YOUR ANSWER HERE:

## What Happens Without the Weird Cases?

Remove the weird observations and compare the two essay conditions again.
Report the number of observations, missing outcomes, and average cleaning score.

Return to the four-cell comparison. Add a `filter()` before `group_by()` and
then group only by `Dissonance`.

Reference for the order of the steps:

```r
example_data |>
  filter(variable_to_keep == value_to_keep) |>
  group_by(group_variable) |>
  summarize(...)
```

```{r comparison-without-weird}
# INSERT CODE HERE
```

Does a difference remain? How does it compare with the pattern created by the
weird cases? What can we conclude from a comparison of means alone at this
point in the course?

YOUR ANSWER HERE:

# What Did We Discover?

Only now should we compare our investigation with the work done by Data Colada.

Uri Simonsohn, Joseph Simmons, and Leif Nelson identified the same twenty
“Harvard” observations. Those observations line up extraordinarily well with
the paper’s predicted result. The article has since been retracted.

Return to the claim at the beginning of the workshop. Do the posted data
provide credible evidence that inauthenticity increases the desire to clean?

YOUR ANSWER HERE:

## What Did the Code Help Us Do?

- `count()` exposed unusual raw responses.
- `mutate()` and `if_else()` turned judgments into variables that could be
  examined.
- Variable names forced us to confront the difference between description and
  interpretation.
- A missing value produced `NA` and forced us to inspect rather than guess.
- `group_by()` and `summarize()` made the decisive four-way comparison visible.
- The code provides a reproducible record of every choice we made.

The point was not simply to obtain the final four means. The path to those
means showed us why the finding cannot be taken at face value.

# New R Functions Introduced in Lecture 4

This workshop primarily uses the cumulative toolkit from Lectures 2 and 3.

| Function | What it does |
|---|---|
| `is.na()` | Identifies values that are missing. |
| `group_by()` | Defines the groups used by a following summary. |
| `n()` | Counts observations within `summarize()`. |
| `sum()` | Adds values; here it counts values for which `is.na()` is true. |

# Appendix: Experiment 4 Materials

The authors' complete posted protocol contains materials for all five
experiments. The portions relevant to our analysis are summarized below. See
the [Materials and Methods](readings/Materials_and_methods.docx) for the full
wording.

## Participants and Initial Questions

Participants had to report that they were Harvard undergraduates. Before the
writing task, they reported:

- whether they supported including course-difficulty ratings in Harvard's Q
  Guide;
- how strongly they held that opinion;
- their age and gender; and
- their year in school.

## Writing Conditions

Participants were assigned to one of three conditions:

1. **Low-choice counterattitudinal:** Participants were assigned to argue
   against their stated position.
2. **High-choice counterattitudinal:** Participants were asked to argue against
   their stated position while the instructions emphasized that doing so was
   voluntary.
3. **High-choice proattitudinal:** Participants were asked to argue in favor of
   their stated position while the instructions emphasized that doing so was
   voluntary.

All participants were told to make the essay persuasive and convincing.

## Cleaning-Product Outcome

After writing the essay, participants rated ten products from 1, “Completely
undesirable,” to 7, “Completely desirable.” Five were cleaning products—Dove
shower soap, Crest toothpaste, Windex cleaner, Tide detergent, and Lysol
disinfectant—and five were filler products. The variable
`av_products_clean` averages the five cleaning-product ratings.

## Post-Treatment Questions

Participants then rated whether the writing task made them feel out of touch
with their “real” selves and how much choice they believed they had in writing
the essay.

# Works Cited

## Data and Study Materials

Gino, Francesca, Maryam Kouchaki, and Adam D. Galinsky. 2015. “The Moral Virtue
of Authenticity: How Inauthenticity Produces Feelings of Immorality and
Impurity.” *Psychological Science* 26: 983–996. Retracted.
[https://doi.org/10.1177/0956797615575277](https://doi.org/10.1177/0956797615575277).

Gino, Francesca, Maryam Kouchaki, and Adam D. Galinsky. “Materials, Methods, and
Posted Data for ‘The Moral Virtue of Authenticity.’” Open Science Framework.
[https://osf.io/sd76g/](https://osf.io/sd76g/). The course file
`GKG_Study4_Class.csv` is a teaching copy of the posted
Study 4 CSV. The only transformation is removal of trailing white space from
`yearSchool`; the preserved source is `GKG_Study4.csv`.

## Writing

Simonsohn, Uri, Joseph P. Simmons, and Leif D. Nelson. 2023. “Data Falsificada
(Part 2): ‘My Class Year Is Harvard.’” *Data Colada* 110, June 20.
[https://datacolada.org/110](https://datacolada.org/110).
