---
title: "Session 6. Uncertainty: Is the Difference Real?"
subtitle: "Polling, the Gender Gap, and Red Versus Blue"
author: "Prof. Josh Clinton"
institute: "Vanderbilt University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Learning Objectives

- Explain the intuition behind a poll's margin of error.

- Explain why a bootstrap samples with replacement.

- Use a loop to repeat the same analysis many times.

- Use `bind_rows()` to collect the result from each loop iteration.

- Use resampling to see how much a percentage or difference of means might vary.

- Use a bootstrap interval to ask whether a difference is distinguishable from zero.

# A Familiar Number: The Margin of Error

A poll might report 52 percent support with a margin of error of plus or minus
3 percentage points. What does that margin of error mean? Why is there a margin
of error if the poll has already counted every answer in its sample?

The percentage is exact for the people who answered the poll. But those people
are only one sample. A different sample would probably produce a somewhat
different percentage.

The margin of error describes this **sampling uncertainty**.

We cannot repeatedly field the same survey. But we can use the sample we have
to imitate that thought experiment:

1. Draw a new sample from the observed sample.
2. Calculate the estimate in that new sample.
3. Repeat this many times.
4. Examine how much the estimates vary.

That procedure is called **resampling**. We will begin with one polling
percentage, then return to the gender gap from Lecture 3.

```{r libraries}
library(tidyverse)
```

# Return to the 2024 CES

In Lecture 3, we created an analysis-ready row-level tibble for the gender gap
and saved it as an `.rds` file. We can read that R object back into memory
without repeating the wrangling.

```{r load-ces}
ces_gender = readRDS("CES2024_GenderGap.rds")

glimpse(ces_gender)
```

`HarrisVoter` equals 1 for a respondent who reported voting for Harris and 0
for a respondent who reported voting for Trump. What does its mean represent?

Because the variable contains only zeros and ones, its mean is the proportion
who reported voting for Harris.

```{r ces-group-means}
ces_gender |>
  summarize(
    N = n(),
    PctHarris = mean(HarrisVoter)
  )
```

That is the exact percentage in this CES teaching sample. How much might it
change if the CES had interviewed a different sample?

# Samples, Populations, and Uncertainty

The **sample** is the data we have: the people interviewed for the CES. The
**population** is the larger group we want to learn about: voters beyond the
particular people who answered this survey.

A sample estimate will usually not be exactly equal to the population value.
Sampling uncertainty is the variation created by observing one sample rather
than another.

Resampling uses the variation in the observed data to approximate what we might
have found in other samples.

# How Does `slice_sample()` Work?

We can use `slice_sample()` to randomly draw rows from a data frame. To make
each part of the function visible, begin with five CES respondents and add an
identification number.

```{r slice-sample-demo-data}
ces_demo = ces_gender |>
  select(gender4, presvote) |>
  head(5) |>
  mutate(case_id = 1:n())

ces_demo
```

The argument `n` tells `slice_sample()` how many rows to draw.

```{r slice-sample-n}
ces_demo |>
  slice_sample(n = 3)
```

By default, the same row cannot be selected twice. Once a row is drawn, it is
temporarily set aside. This is **sampling without replacement**.

Instead of specifying the number of rows, `prop` tells `slice_sample()` what
**proportion** of the rows to draw. Here `.4` means 40 percent of the five rows,
so the result contains two rows.

```{r slice-sample-prop}
ces_demo |>
  slice_sample(prop = .4)
```

Suppose we want an imaginary new poll with the same number of respondents as
the observed poll. We might begin with `prop = 1`.

```{r missing-replacement}
ces_demo |>
  slice_sample(prop = 1)
```

The code runs, but look at `case_id`. We sampled all five rows without
replacement, so R had no choice but to return every row exactly once. It only
changed their order. Would rearranging the respondents change their mean?

This is a useful error: the function is correct, but a critical argument is
missing.

For a bootstrap sample, we use `prop = 1` because the new sample should be the
same size as the original data. We also use `replace = TRUE`. After a row is
drawn, it is put back and can be drawn again.

```{r slice-sample-replacement}
set.seed(1)

ces_demo |>
  slice_sample(prop = 1, replace = TRUE)
```

Look at `case_id`. Some respondents appear more than once and others do not
appear at all. The bootstrap sample still contains five rows, but it is not
identical to the original five rows.

The complete instruction therefore means:

- `prop = 1`: draw 100 percent as many rows as are in the original data;
- `replace = TRUE`: put each selected row back so it can be selected again.

# One Imaginary New Poll

Now use those instructions with the entire CES dataset. Because we want our
work to be reproducible, we set a seed before drawing.

```{r one-ces-resample}
set.seed(42)

ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))
```

Run the code several times without resetting the seed. The estimate changes
because each resample contains a somewhat different collection of respondents.

One resample shows what could happen once. It does not tell us how much the
estimate usually changes. We need to repeat the same analysis many times.

# What Does a Loop Do?

A loop tells R to repeat a block of code. Begin with a loop that does nothing
except print the value of `i`.

```{r loop-counts}
for (i in 1:5) {
  print(i)
}
```

Read the first line as: “For each value of `i` from 1 through 5, run the code
inside the braces.”

The braces identify what is repeated. On the first pass, `i` equals 1. On the
second pass, `i` equals 2. The loop stops after the fifth pass.

Now replace `print(i)` with the work we want to repeat:

1. Draw a bootstrap sample.
2. Calculate Harris support.
3. Store the answer.

# Do the Work Once Before Repeating It

Before writing the loop, do the work once.

```{r one-harris-estimate}
one_estimate = ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))

one_estimate
```

`one_estimate` is a one-row tibble. If we repeat this step, we need an object
that can collect each one-row answer.

# What Does `bind_rows()` Do?

Begin with an empty object.

```{r empty-ces-container}
SampledHarrisSupport = NULL
```

`bind_rows()` places the rows from one object underneath the rows in another
object. Here the first object is empty, so the result contains the one row from
`one_estimate`.

```{r bind-first-result}
SampledHarrisSupport = bind_rows(
  SampledHarrisSupport,
  one_estimate
)

SampledHarrisSupport
```

Now make another estimate and bind it to the object that already contains the
first estimate.

```{r bind-second-result}
another_estimate = ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))

SampledHarrisSupport = bind_rows(
  SampledHarrisSupport,
  another_estimate
)

SampledHarrisSupport
```

Why are there now two rows? `bind_rows()` does not replace the first estimate.
It adds the second estimate beneath it.

# Put the Steps Inside a Loop

Start with five iterations so we can inspect the result.

```{r five-harris-resamples}
SampledHarrisSupport = NULL
set.seed(2300)

for (i in 1:5) {

  one_estimate = ces_gender |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Iteration = i,
      PctHarris = mean(HarrisVoter)
    )

  SampledHarrisSupport = bind_rows(
    SampledHarrisSupport,
    one_estimate
  )
}

SampledHarrisSupport
```

The loop repeats the same commands five times. `bind_rows()` stores one answer
on each pass. The value of `i` lets us see which pass produced each row.

What should change if we replace `1:5` with `1:1000`? What should remain the
same?

# Repeat the Poll 1,000 Times

Once the five-iteration loop works, change only the number of repetitions.

```{r harris-bootstrap-loop}
SampledHarrisSupport = NULL
set.seed(2300)

for (i in 1:1000) {

  one_estimate = ces_gender |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(PctHarris = mean(HarrisVoter))

  SampledHarrisSupport = bind_rows(
    SampledHarrisSupport,
    one_estimate
  )
}
```

Each row now contains Harris support from one imaginary new poll.

```{r inspect-ces-bootstrap}
summary(SampledHarrisSupport)
```

The `summary()` output describes the **bootstrap distribution** numerically.
Compare its median to Harris support in the original sample. The minimum,
quartiles, and maximum show how much the estimate changes across resamples.

# From a Margin of Error to an Interval

We can use the 2.5th and 97.5th percentiles to identify the middle 95 percent
of our resampled estimates.

```{r ces-bootstrap-interval}
SampledHarrisSupport |>
  summarize(
    Lower = quantile(PctHarris, .025),
    Estimate = mean(PctHarris),
    Upper = quantile(PctHarris, .975),
    MarginOfError = (Upper - Lower) / 2
  )
```

This is a **95 percent bootstrap interval**. Its margin of error is half the
width of the interval.

Does a narrow interval mean the CES estimate must match the election result?
No. The bootstrap cannot add people who were systematically missing from the
survey or correct a sample that overrepresents some kinds of respondents.
The margin of error describes sampling error, not every possible source of
polling error.

# A Harder Question: Is the Gender Gap "Real"?

In Lecture 3, we found different levels of Harris support among women and men.
Calculate the two group means again.

```{r gender-group-means}
ces_gender |>
  group_by(female) |>
  summarize(
    N = n(),
    PctHarris = mean(HarrisVoter)
  )
```

The statistic is now a **difference of means**: Harris support among women
minus Harris support among men.

```{r observed-gender-gap}
Women = ces_gender |>
  filter(female == 1) |>
  summarize(PctHarris = mean(HarrisVoter))

Men = ces_gender |>
  filter(female == 0) |>
  summarize(PctHarris = mean(HarrisVoter))

GenderGap = Women$PctHarris - Men$PctHarris
GenderGap
```

Women in this sample were about `r round(100 * GenderGap, 1)` percentage points
more likely than men to report voting for Harris. That describes this sample.
How much might the estimated gap change in a different sample?

# Bootstrap the Gender Gap

The loop has the same three parts: sample, calculate, store. Only the statistic
inside the loop changes.

```{r gender-gap-bootstrap-loop}
SampledGenderGaps = NULL
set.seed(2300)

for (i in 1:1000) {

  one_sample = ces_gender |>
    slice_sample(prop = 1, replace = TRUE)

  one_women = one_sample |>
    filter(female == 1) |>
    summarize(PctHarris = mean(HarrisVoter))

  one_men = one_sample |>
    filter(female == 0) |>
    summarize(PctHarris = mean(HarrisVoter))

  one_gap = tibble(
    GenderGap = one_women$PctHarris - one_men$PctHarris
  )

  SampledGenderGaps = bind_rows(SampledGenderGaps, one_gap)
}
```

```{r inspect-gender-gap-bootstrap}
summary(SampledGenderGaps)
```

```{r gender-gap-bootstrap-interval}
SampledGenderGaps |>
  summarize(
    Lower = quantile(GenderGap, .025),
    Estimate = mean(GenderGap),
    Upper = quantile(GenderGap, .975)
  )
```

Does the interval include zero? Here the interval is above zero. The gender gap
is distinguishable from zero as sampling error.

Does that mean the gender gap is "real"? We need to be precise:

- The difference is too large to explain easily as sampling variation alone.
- It remains a descriptive association, not evidence that gender caused vote
  choice.
- The bootstrap does not fix people being missing or overrepresented in the
  survey. This teaching extract is unweighted, and we already know its raw vote
  estimate does not match the election result.

The margin of error describes **sampling error**. It does not describe every
possible source of error in a poll.

# Same Method, New Question

The bootstrap is useful because it is not limited to polling. We can use the
same procedure for any statistic we can calculate in a sample.

At the 2004 Athens Olympics, competitors in boxing, taekwondo, Greco-Roman
wrestling, and freestyle wrestling were assigned red or blue outfits. Hill and
Barton reported that red won significantly more often than blue in the pooled
data ($p=.041$). They argued that wearing red might influence the outcome of a
close contest.^[Russell A. Hill and Robert A. Barton. 2005. "[Red Enhances Human Performance in Contests](https://www.nature.com/articles/435293a)."
*Nature* 435: 293.]

Before opening the data, what is the question?

- What is the unit being studied?
- What is assigned to red or blue: the athlete or the position in a bout?
- What is the outcome?
- What comparison would show a red advantage?
- Why does random assignment matter?
- What would it mean for the difference to equal zero?

The basic comparison is the rate at which red wins minus the rate at which blue
wins. If color makes no difference, that difference should be zero. A positive
difference means red won more often; a negative difference means blue won more
often.

The questions are now familiar:

- How large was the red-blue difference?
- How much might that estimate vary in a different sample of bouts?
- Is the difference distinguishable from zero?

```{r load-athens}
athens = read_csv("athens.csv") |>
  mutate(
    red_win = if_else(winner == "Red", 1, 0),
    blue_win = if_else(winner == "Blue", 1, 0)
)
```

Each row is one bout. `red_win` equals 1 when red won and 0 when red lost.
`blue_win` does the reverse. Before running the next chunk, what should the two
means add to? Why?

# The Red-Blue Difference of Means

The mean of each indicator is the win rate for that color. 
Calculate the two means and subtract.

```{r observed-red-blue}
athens |>
  summarize(
    RedWinRate = mean(red_win),
    BlueWinRate = mean(blue_win),
    Difference = RedWinRate - BlueWinRate
  )
```

The estimated difference is about 8.5 percentage points. Red won more often in
these data. Now apply the procedure we just learned.

# Bootstrap the Red-Blue Difference

```{r red-blue-bootstrap}
SampledDifferences = NULL
set.seed(2300)

for (i in 1:1000) {

  one_difference = athens |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(blue_win)
    )

  SampledDifferences = bind_rows(
    SampledDifferences,
    one_difference
  )
}
```

```{r inspect-red-blue-bootstrap}
summary(SampledDifferences)
```

Compare this summary to the CES bootstrap. Which set of estimates varies more
from one resample to another?

```{r red-blue-interval}
SampledDifferences |>
  summarize(
    Lower = quantile(Difference, .025),
    Estimate = mean(Difference),
    Upper = quantile(Difference, .975)
  )
```

Does this interval include zero? Is the Olympic estimate more or less precise
than the CES gender-gap estimate? Why?

Hill and Barton reported a statistically significant pooled result
($p=.041$). Our simple difference using every bout in their posted supplement
is less certain: its interval includes zero. This does not by itself prove that
the published result is wrong. It tells us that the answer depends on exactly
which comparison and analysis are being used. Their paper also examines
rounds, weight classes, and contests between athletes of similar ability.

That creates another question: which analysis was specified because it follows
from the argument, and which analyses became important after looking at the
data? We cannot answer that from the p-value alone.

# Testing the Difference Against Zero

A confidence interval asks whether zero is a plausible value for the population
difference. A hypothesis test asks a closely related question:

> If red and blue were equally likely to win, how often would chance produce a
> difference at least as large as the one we observed?

Under this **null hypothesis**, each bout is like a fair coin flip: red has a 50 percent chance of winning and blue has a 50 percent chance.

```{r observed-difference-object}
ObservedDifference = athens |>
  summarize(Difference = mean(red_win) - mean(blue_win)) |>
  pull(Difference)
```

Simulate 1,000 imaginary Olympics in which color has no advantage.

```{r null-loop}
NullDifferences = NULL
fair_coin = tibble(red_win = c(0, 1))
set.seed(2300)

for (i in 1:1000) {

  one_null = fair_coin |>
    slice_sample(n = nrow(athens), replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(1 - red_win)
    )

  NullDifferences = bind_rows(NullDifferences, one_null)
}
```

```{r inspect-null}
summary(NullDifferences)
```

The null differences are centered near zero because they come from a world in
which red and blue are equally likely to win. We now count how many imaginary
Olympics produced a difference at least as far from zero as the observed
difference, in either direction.

```{r p-value}
NullDifferences |>
  summarize(
    p_value = mean(abs(Difference) >= abs(ObservedDifference))
  )
```

This simulated **p-value** is the fraction of fair-coin results at least as extreme as the observed result. 
A small p-value is evidence against the no-difference hypothesis. 
It is not the probability that the null hypothesis is true.

# What Happens When We Add More Data?

The Athens result comes from one Olympic Games. 
We now have comparable results from Olympics held between 1996 and 2020.

```{r load-all-olympics}
olympics = read_csv(
  "olympics.csv"
) |>
  mutate(
    red_win = if_else(winner == "Red", 1, 0),
    blue_win = if_else(winner == "Blue", 1, 0)
  )
```

Calculate the difference for each Olympic year.

```{r differences-by-year}
olympics |>
  group_by(year) |>
  summarize(
    N = n(),
    Difference = mean(red_win) - mean(blue_win)
  )
```

The result is not stable across Olympics. In some years red wins more; in
others, blue wins more.

Now calculate the difference using all of the contests.

```{r pooled-difference}
olympics |>
  summarize(
    RedWinRate = mean(red_win),
    BlueWinRate = mean(blue_win),
    Difference = RedWinRate - BlueWinRate
  )
```

The apparent red advantage is much smaller when we include more data. Use the
same bootstrap one more time.

```{r all-years-bootstrap}
AllYearsDifferences = NULL
set.seed(2300)

for (i in 1:1000) {

  one_difference = olympics |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(blue_win)
    )

  AllYearsDifferences = bind_rows(
    AllYearsDifferences,
    one_difference
  )
}

AllYearsDifferences |>
  summarize(
    Lower = quantile(Difference, .025),
    Estimate = mean(Difference),
    Upper = quantile(Difference, .975)
  )
```

Does this interval include zero? What has changed since we looked only at
Athens in 2004?

# What Have We Learned?

- A margin of error describes how much an estimate might change across samples.
- Bootstrap resampling imitates repeated samples using the data we have.
- The same procedure works for a poll percentage or a difference of means.
- A confidence interval helps us ask whether zero is plausible.
- A hypothesis test asks how surprising our estimate would be if the true
  difference were zero.
- Sampling uncertainty is only one kind of uncertainty; resampling cannot fix
  a sample that poorly represents the population.
- A finding from one dataset can weaken when more evidence becomes available.

# New R Functions Introduced in Lecture 6

Functions introduced in earlier lectures are intentionally omitted.

| Function | What it does |
|---|---|
| `slice_sample()` | Randomly draws rows from a data frame. |
| `set.seed()` | Makes random results reproducible. |
| `tibble()` | Creates a data frame in tidyverse format. |
| `bind_rows()` | Adds rows to the bottom of a data frame. |
| `quantile()` | Returns values at specified percentiles. |
| `pull()` | Extracts one variable from a data frame as a vector. |
| `abs()` | Returns the absolute distance from zero. |
| `nrow()` | Returns the number of rows in a data frame. |
| `for` | Repeats a block of code a specified number of times. |

# Works Cited

Hill, Russell A., and Robert A. Barton. 2005. “Red Enhances Human Performance
in Contests.” *Nature* 435: 293.
[https://doi.org/10.1038/435293a](https://doi.org/10.1038/435293a).

Peperkoorn, Leonard S., Russell A. Hill, Robert A. Barton, and Thomas V. Pollet.
2024. “Meta-analysis of the Red Advantage in Combat Sports.” *Scientific
Reports* 14: 30822.
[https://doi.org/10.1038/s41598-024-81373-3](https://doi.org/10.1038/s41598-024-81373-3).
Data and code: [https://osf.io/fqkcw/](https://osf.io/fqkcw/).

Schaffner, Brian, Marissa Shih, Stephen Ansolabehere, and Jeremy Pope. 2025.
“Cooperative Election Study Common Content, 2024.” Harvard Dataverse, V9.
[https://doi.org/10.7910/DVN/X11EP6](https://doi.org/10.7910/DVN/X11EP6).
