---
title: "Session 6. Uncertainty: Is the Difference Real?"
subtitle: "Polling, the Gender Gap, and Red Versus Blue"
author: "Prof. Josh Clinton"
institute: "Vanderbilt University"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Learning Objectives

- Explain the intuition behind a poll's margin of error.

- Explain why a bootstrap samples with replacement.

- Use a loop to repeat the same analysis many times.

- Use `bind_rows()` to collect the result from each loop iteration.

- Use resampling to see how much a percentage or difference of means might vary.

- Use a bootstrap interval to ask whether a difference is distinguishable from zero.

## Uncertainty in Univariate Statistics

When we calculate a summary statistic in univariate statistics, we're making a statement that can be interpreted in terms of what we can expect to see in other similar situations. If I say that the average height of a cedar tree is 75 feet, for example, that provides an expectation -- prediction -- for what we might expect to find if we look at other cedar trees (assuming "similar" environmental conditions). The measure we propose to measure what is typical is almost naturally interpreted by humans in terms of a prediction for what we may expect of other similar units.^[As we shall see in this course, the word "similar" here is doing a lot of work! Critical to this notion of prediction is the idea that the data we are analyzing "looks like" the data we are trying to predict!]

However, there's more information that we need to communicate beyond what is "typical." We also care about our level of uncertainty around that measure. Sure, the average height might be 75 feet, but does that mean in every sample we ever collect we're always going to see an average of 75 feet?  How likely is it that we may expect a tree that is much larger or much smaller?  There are many ways to get an average of 75 - perhaps every single tree is exactly 75 feet.  Or perhaps half of the trees are 149 feet tall and the other half are only 1 foot tall.  Whether we are in case 1 or case 2 likely matters a lot!

Today we will discuss how we can leverage the variation we observe in the data that we have to quantify how much confidence we should have in the precision/certainty/variation of our measures. 

# A Familiar Number: The Margin of Error

A poll might report 52 percent support with a margin of error of plus or minus 3 percentage points. What does that margin of error mean? Why is there a margin of error if the poll has already counted every answer in its sample?

The tools that we are learning can be thought of in two ways.  One way is that they are a means to analyze the data that we have and to reach conclusions about what we have observed.  All we wish to do is to characterize the data we have collected.  This is the task of _description_.  We are not trying to use the data for anything beyond talking about the observed data.  In the case of the Exit Polls we looked at, this would mean being content with studying only the opinions of those who were actually interviewed.

But frequently we want to use the data to learn about the larger world. We want to talk about cases that have not been measured.  In other words, we want to use the data to make an interpretation about what the variation in the data reveals about variation in cases that are not contained in our data. This is typically what we are trying to do -- we are doing an Exit Poll not to talk about the people that were interviewed, but rather to talk about what voters in general are thinking and doing.  This can involve the task of _description_ (e.g., what fraction of the electorate has a post-graduate degree, is over the age of 65, etc) or it can involve the task of _prediction_ (e.g., who is likely to win the state given the opinions revealed in the Exit Poll).  We can also use data for the _evaluation_ of claims about what is true in the world -- i.e., determine whether certain claims about the world are true (e.g, are athletes who wear the color Red in competitions more likely to beat athletes who wear the color Blue?)

Whenever we want to use the data to learn about data that we have not collected, it becomes critically important to consider how the data is collected. We often want to jump right in and start analyzing, but thinking about where the data comes from is an essential first step if we wish to use the data to learn about cases outside of our data.

To help, let’s define two terms: *Population* and *Sample*.

An *observation* is simply a data point in our *sample*. It is what a row in our data represents.  Each observation is associated with a bunch of characteristics (variables).

A *Sample* is the data that you are able to analyze. It could be a few observations or hundreds of millions. Somewhat confusingly, it does not have to be gathered by intentionally sampling anything. If you were to write down the temperature each day for 100 days, that would be a sample of temperature data. If I were to write some public opinion questions and then ask them to the first ten people I met to record their responses that would also be a sample of public opinion data. Not a very good sample, but a sample nonetheless. Whatever data you are able to analyze is a sample – think about sample in this context as meaning simply the data you have available to analyze.

The *Population* is the set of observations to which the sample belongs.what you want to learn about. This is where things get a bit tricky and confusing. In statistics, we can never observe population – it is a hypothetical construct about the observations of interest because it includes cases that have yet to happen. Seem weird? Let’s try to explain this. Suppose that you collected data on the national popular vote for every presidential election in the United States to the current day. Your sample therefore contains the results of every single presidential election contest. There are no ``secret" elections and you did not forget to include any. You have it all. Despite the fact that your sample contains every presidential election, we often want to make statements about what we expect the results to be based on the sample of data. Put differently, if you use the data to determine how much the winner defeats the loser in terms of percentage points then we are using the data to make a statement not only about what the margin has historically been on average, but also what we would expect the margin to be in the future (assuming the political context is relatively stable).1

The important distinction is that sample refers to the _observations in the data you are analyzing_ and population refers to the _observations that you are trying to learn_ about using the sample. Note that if we survey the entire population -- i.e., our sample is the population -- then we call this a **Census**.

The field of statistics is all about how we can try to use samples to learn – that is to make inferences – about the population. Under what conditions are we able to say something? How likely is it that our estimates are correct? How certain should we that our estimates are close to the truth? And what do we mean by ``close to the truth"?

We are going to focus on a particular approach to quantifying our ability to learn that uses the data. While useful for explaining the concepts because it uses the data and our programming skills to directly illustrate the underlying logic, it is not the most efficient way to do so – especially when working with large number of observations where this can be computationally infeasible. (This is where math comes in and the ability to use math to derive the results analytically instead of relying on computational power.

Critical to this approach -- and indeed most of what we do in statistics -- is the assumption that _our sample is a random sample of the population_.^[Technically speaking, a random sample means that every unit in the population has an equal chance of being included in the data.  In the polling world, this means that every registered voter has the same opporuntity of being selected for a poll.  (Reality is more complicated than this as we don't actually need to have equal probabilities of selection so long as we know what the probability of selection is, but for this class let's keeep the simpler definition and assumption!).]  If so, the sample of data that we collect can be used to summarize the population data that we are unable to collect!  In other words, we can use the results of polling 1000 people to summarize the opinions of a much larger polity.

So a **sample** is a subset of the **population** and, because of this, the estimate that we get from a sample will not usually be exactly equal to the population value. A **statistic** is what we call a function of the sample -- hence the study of statistics is the study of how we use the analysis of a sample (by applying functions to sample data) to learn about what is true in the population.  But because we observe only some of the population -- a sample -- we need to account for the fact that we have some uncertainty about how close the answer we compute using our sample is to the actual truth in the population.  Put differently, sampling variability is the variation that results from using and analyzing one sample rather than another.


Put differently, we can exactly calculate the results for people who answered the poll. But those people are only one sample out of many and a different sample would probably produce a somewhat different percentage.

The margin of error describes this **sampling uncertainty**.

We cannot repeatedly field the same survey. But we can use the sample we have
to imitate that thought experiment:

1. Draw a new sample from the observed sample.
2. Calculate the estimate in that new sample.
3. Repeat this many times.
4. Examine how much the estimates vary.

That procedure is called **resampling**. We will begin with one polling
percentage, then return to the gender gap from Lecture 3.

```{r libraries}
library(tidyverse)
```

# Return to the 2024 CES

Let us return to the data we were using in Session 3. We are going to load an analysis-ready dataset related to measuring the gender gap that was saved as an `.rds` file. (This is a file format that is unique to R rather than a generic file format like the comma-separated-values (csv) files that we have used previously.  We can read this into memory using the `readRDS` function and then take a glimpse.
without repeating the wrangling.

```{r load-ces}
ces_gender = readRDS("CES2024_GenderGap.rds")

glimpse(ces_gender)
```

As before, `HarrisVoter` equals 1 for a respondent who reported voting for Harris and 0 for a respondent who reported voting for Trump. Because this is a binary variable, what does its mean represent?

Because the variable contains only zeros and ones - sometimes called an "Indicator" (or "Dummy") variable, its mean is the proportion of respondents who reported voting for Harris.  Let's compute that.

```{r ces-group-means}
ces_gender |>
  summarize(
    N = n(),
    PctHarris = mean(HarrisVoter)
  )
```

This is the exact percentage that we observe in this specific sample of survey respondents. But much might it change if the CES had interviewed a different sample?  If we treat respondents as a random sample of the population, the amount that the answers differs due to randomness alone is called the ``margin of error."  

Note that this is not the same as survey error!  The margin of error is how much a survey-estimate may change as a result of random sampling and the fact that some people were randomly selected to be interviewed while others were not.

# Samples, Populations, and Uncertainty

We're going to start by building up a range of uncertainty from the data we already have. We'll do this by sampling from the data itself. 

We are going to use the computer to construct random samples. Of course, the numbers are not really random because they are computers. The upside of this is that we can ensure replicability by choosing a particular seed value. To do so we are going to use the function `set.seed`.

By setting a specific seed we are ensuring that anyone who sets the same seed value will get exactly the same set of random numbers. This is super useful for debugging and replicability as it allows us to remove the impact of the random number generation (which isn't really truly random).

For my seed value I am going to choose the answer to life, the universe, and everything. (Does anyone get this reference or am I officially too old?). 

*NOTE: Always set a seed if you are sampling so you can replicate your results!*

```{r}
set.seed(42)
```


Resampling uses the variation in the observed data to approximate what we might have found in other samples.

# How Does `slice_sample()` Work?

We can use `slice_sample()` to randomly draw rows from a data frame. To make each part of the function visible, begin with the first five CES respondents -- which we extract using the `head` function -- and add an identification number.

```{r slice-sample-demo-data}
ces_demo = ces_gender |>
  select(gender4, presvote) |>
  head(5) |>
  mutate(case_id = 1:n())

ces_demo
```

The argument `n` tells `slice_sample()` how many rows to draw.

```{r slice-sample-n}
ces_demo |>
  slice_sample(n = 3)
```

By default, the same row cannot be selected twice. Once a row is drawn, it is temporarily set aside. This is **sampling without replacement**.

Instead of specifying the number of rows, `prop` tells `slice_sample()` what **proportion** of the rows to draw. Here `.4` means 40 percent of the five rows, so the result contains two rows.

```{r slice-sample-prop}
ces_demo |>
  slice_sample(prop = .4)
```

Suppose we want an imaginary new poll with the same number of respondents as the observed poll. We might begin with `prop = 1`.

```{r missing-replacement}
ces_demo |>
  slice_sample(prop = 1)
```

The code runs, but look at `case_id`. We sampled all five rows without replacement, so R had no choice but to return every row exactly once. It only changed their order. Would rearranging the respondents change their mean?

What we have done is to _sample without replacement_. What this means is that once we sample/select an observation, we remove it and sample/select from the remaining observations. If we sample the same number of observations without replacement, the result is a random rearrangment of the original object. While this can be useful in some contexts, this is not what we want here. 

The function is correct, but a critical argument is missing. We need to specify that we want _sample with replacement_. 

For a bootstrap sample, we use `prop = 1` because the new sample should be the same size as the original data. We also use `replace = TRUE` -- this means that after a row is drawn, it is put back and can be drawn again.

```{r slice-sample-replacement}
set.seed(42)

ces_demo |>
  slice_sample(prop = 1, replace = TRUE)
```

Look at `case_id`. Some respondents appear more than once and others do not appear at all. The bootstrap sample still contains five rows, but it is no longer identical to the original five rows.

The complete function therefore reads:

- `prop = 1`: draw 100 percent as many rows as are in the original data;
- `replace = TRUE`: put each selected row back so it can be selected again.

# One Imaginary New Poll

Now use those instructions with the entire CES dataset. Because we want our work to be reproducible, we `set.seed` before drawing.  If you do not do this you get different answers.

```{r one-ces-resample}
set.seed(42)

ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))
```

One resample shows what could happen once. But it does not tell us how much the estimate may change if we were to do this over and over again. To do this we need to repeat the same analysis many times.

# What Does a Loop Do?

A loop tells R to repeat a block of code for as many times as we tell it -- here `5`. Note the syntax and the braces. Code within the braces is what R will do each time -- here nothing but print the value of `i`.

```{r loop-counts}
for (i in 1:5) {
  print(i)
}
```

Read the first line as: “For each value of `i` from 1 through 5, run the code inside the braces.”

On the first pass, `i` equals 1. On the second pass, `i` equals 2. The loop stops after the fifth pass when `i==5`.

Of course we would not want to do this for real (although printing `i` is sometimes good when debugging as it shows when the code breaks). Now replace `print(i)` with the work we want to repeat:

1. Draw a bootstrap sample.
2. Calculate Harris support.
3. Store the answer.

# Do the Work Once Before Repeating It

Before writing the loop, do the work once to make sure it works!  Once we confirm that it works then we can build the loop.

```{r one-harris-estimate}
one_estimate = ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))

one_estimate
```

`one_estimate` is a one-row tibble. If we repeat this step, we need an object that can collect each one-row answer.  Otherwise we will end up overwriting what we just did!

Note also that we are not saving our resampled data - we are sampling, summarizing, and forgetting.  The reason is that it we want to do lots of samples and/or we have a large number of observations being sampled then storing all of those random samples can impact computing performance by keeping large objects in memory. As a result, we generally try to draw a sample, perform the analysis that we need on the sample that was selected, and save only the most essential results of that analysis.  (What would we change to save the resampled data?)

# What Does `bind_rows()` Do?

Begin with an empty object - which we define to be `NULL` to start.

```{r empty-ces-container}
SampledHarrisSupport = NULL
```

The `bind_rows()` function places the rows from one object underneath the rows of another object. Because the first object here is empty, the result contains the one row from `one_estimate`.

```{r bind-first-result}
SampledHarrisSupport = bind_rows(
  SampledHarrisSupport,
  one_estimate
)

SampledHarrisSupport
```

Now let's make another estimate and bind it to the object that already contains the first estimate.

```{r bind-second-result}
another_estimate = ces_gender |>
  slice_sample(prop = 1, replace = TRUE) |>
  summarize(PctHarris = mean(HarrisVoter))

SampledHarrisSupport = bind_rows(
  SampledHarrisSupport,
  another_estimate
)

SampledHarrisSupport
```

Why are there now two rows? `bind_rows()` does not replace the first estimate; it adds the second estimate beneath it.  This is looking promising!

# Put the Steps Inside a Loop

Now let's pull this all together.  Having shown that the code we want works, lets put the code in a loop. Let's start with five iterations so we can inspect the result.

```{r five-harris-resamples}
SampledHarrisSupport = NULL
set.seed(42)

for (i in 1:5) {

  one_estimate = ces_gender |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Iteration = i,
      PctHarris = mean(HarrisVoter)
    )

  SampledHarrisSupport = bind_rows(
    SampledHarrisSupport,
    one_estimate
  )
}

SampledHarrisSupport
```

The loop repeats the same commands five times. `bind_rows()` stores one answer on each pass. The value of `i` lets us see which iteration produced each estimate.

Now think about what happens if we replace `1:5` with `1:1000`? What should remain the same?

# Repeat the Poll 1,000 Times

Once the five-iteration loop works, change only the number of repetitions.

```{r harris-bootstrap-loop}
SampledHarrisSupport = NULL
set.seed(42)

for (i in 1:1000) {

  one_estimate = ces_gender |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(PctHarris = mean(HarrisVoter))

  SampledHarrisSupport = bind_rows(
    SampledHarrisSupport,
    one_estimate
  )
}
```

Each row now contains Harris support from one imaginary new poll.

```{r inspect-ces-bootstrap}
summary(SampledHarrisSupport)
```

The `summary()` output describes the **bootstrap distribution** numerically.
Compare its median to Harris support in the original sample. The minimum,
quartiles, and maximum show how much the estimate changes across resamples.

# From a Margin of Error to an Interval

We can use the 2.5th and 97.5th percentiles to identify the middle 95 percent of our resampled estimates.

```{r ces-bootstrap-interval}
SampledHarrisSupport |>
  summarize(
    Lower = quantile(PctHarris, .025),
    Estimate = mean(PctHarris),
    Upper = quantile(PctHarris, .975),
    MarginOfError = (Upper - Lower) / 2
  )
```

This is a **95 percent bootstrap interval**. Its margin of error is half the width of the interval.

Does a narrow interval mean the CES estimate must match the election result? No. 

The bootstrap cannot add people who were systematically missing from the survey or correct a sample that overrepresents some kinds of respondents.
The margin of error describes the error that we can attribute to random sampling alone -- not every possible source of polling error.
The fact that the true answer is well-outside of the range that we find suggests that there the polling error is not just because we happened to draw a "bad sample" -- instead, there is something more concerning going on.

# A Harder Question: Is the Gender Gap "Real"?

In Session 3, we found different levels of Harris support among women and men. Let's calculate the two group means again.

```{r gender-group-means}
ces_gender |>
  group_by(female) |>
  summarize(
    N = n(),
    PctHarris = mean(HarrisVoter)
  )
```

The statistic used to measure the gender gap is just a **difference of means**: Harris support among females minus Harris support among males.

Recall how we computed it.

```{r observed-gender-gap}
Female = ces_gender |>
  filter(female == 1) |>
  summarize(PctHarris = mean(HarrisVoter))

Male = ces_gender |>
  filter(female == 0) |>
  summarize(PctHarris = mean(HarrisVoter))

GenderGap = Female$PctHarris - Male$PctHarris
GenderGap
```

Females in this sample were about `r round(100 * GenderGap, 1)` percentage points more likely than males to report voting for Harris. 
But how confident can we be about that estimate?  Might it be different in a different sample?

# Bootstrap the Gender Gap

Let's go!  The loop has the same three parts as before: sample, calculate, store. 
The only thing we need to change is the code within the loop.  That and using the function `tibble` to turn the calculation into an object we can add to `SampledGenderGaps`.

```{r gender-gap-bootstrap-loop}
SampledGenderGaps = NULL
set.seed(42)

for (i in 1:1000) {

  one_sample = ces_gender |>
    slice_sample(prop = 1, replace = TRUE)

  one_female = one_sample |>
    filter(female == 1) |>
    summarize(PctHarris = mean(HarrisVoter))

  one_male = one_sample |>
    filter(female == 0) |>
    summarize(PctHarris = mean(HarrisVoter))

  one_gap = tibble(
    GenderGap = one_female$PctHarris - one_male$PctHarris
  )

  SampledGenderGaps = bind_rows(SampledGenderGaps, one_gap)
}
```

```{r inspect-gender-gap-bootstrap}
summary(SampledGenderGaps)
```

```{r gender-gap-bootstrap-interval}
SampledGenderGaps |>
  summarize(
    Lower = quantile(GenderGap, .025),
    Estimate = mean(GenderGap),
    Upper = quantile(GenderGap, .975)
  )
```

Does the interval include zero? Here the interval is above zero. The gender gap is distinguishable from zero as sampling error.

Does that mean the gender gap is "real"? We need to be precise:

- The difference is too large to explain easily as sampling variation alone.
- It remains a descriptive association, not evidence that gender caused vote
  choice.
- The bootstrap does not fix people being missing or overrepresented in the
  survey. This teaching extract is unweighted, and we already know its raw vote
  estimate does not match the election result.

The margin of error describes **sampling error**. It does not describe every possible source of error in a poll.


# ASIDE: What if our survey was smaller?  

The impact of uncertainty depends on how large our sample is.  To see this, lets repeat this with a smaller sample -- lets just take 1000 observations and pretend that was our actual survey

```{r}
ces_gender_small = ces_gender |>
    slice_sample(n = 1000, replace = FALSE)
```

Now replicate the code from above using this smaller data.

```{r}
# INSERT CODE HERE
```

What do you observe?  Is it different?  How does the size of the interval compare?

# Same Method, New Question

The bootstrap is useful because it is not limited to polling. We can use the same procedure for any statistic we can calculate in a sample!  Even those that have nothing to do with a survey.

At the 2004 Athens Olympics, competitors in boxing, taekwondo, Greco-Roman wrestling, and freestyle wrestling were assigned red or blue outfits. Hill and Barton reported that red won significantly more often than blue in the pooled data ($p=.041$). They argued that wearing red might influence the outcome of a close contest.^[Russell A. Hill and Robert A. Barton. 2005. "[Red Enhances Human Performance in Contests](https://www.nature.com/articles/435293a)." *Nature* 435: 293.]

Before opening the data, what is the question?

- What is the unit being studied?
- What is assigned to red or blue: the athlete or the position in a bout?
- What is the outcome?
- What comparison would show a red advantage?
- Why does random assignment matter?
- What would it mean for the difference to equal zero?

The basic comparison is the rate at which red wins minus the rate at which blue wins. If color makes no difference, that difference should be zero. A positive difference means red won more often; a negative difference means blue won more often.

The questions are now familiar:

- How large was the red-blue difference?
- How much might that estimate vary in a different sample of bouts?
- Is the difference distinguishable from zero?

```{r load-athens}
athens = read_csv("athens.csv") |>
  mutate(
    red_win = if_else(winner == "Red", 1, 0),
    blue_win = if_else(winner == "Blue", 1, 0)
)
```

Each row is one contest. `red_win` equals 1 when red won and 0 when red lost.
`blue_win` does the reverse. Before running the next chunk, what should the two
means add to? Why?

# The Red-Blue Difference of Means

The mean of each indicator is the win rate for that color.  Calculate the two means and subtract.

```{r observed-red-blue}
athens |>
  summarize(
    RedWinRate = mean(red_win),
    BlueWinRate = mean(blue_win),
    Difference = RedWinRate - BlueWinRate
  )
```

The estimated difference is about 8.5 percentage points. Red won more often in these data. Now apply the procedure we just learned.

# Bootstrap the Red-Blue Difference

Now lets resample/bootstrap the analysis.

```{r red-blue-bootstrap}
SampledDifferences = NULL
set.seed(42)

for (i in 1:1000) {

  one_difference = athens |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(blue_win),
      n = n()
    )

  SampledDifferences = bind_rows(
    SampledDifferences,
    one_difference
  )
}
```

```{r inspect-red-blue-bootstrap}
summary(SampledDifferences)
```

Compare this summary to the CES bootstrap. Which set of estimates varies more from one resample to another?

```{r red-blue-interval}
SampledDifferences |>
  summarize(
    Lower = quantile(Difference, .025),
    Estimate = mean(Difference),
    Upper = quantile(Difference, .975)
  )
```

Does this interval include zero? Is the Olympic estimate more or less precise than the CES gender-gap estimate? Why?

Hill and Barton reported a statistically significant pooled result ($p=.041$). Our simple difference using every bout in their posted supplement is less certain and the interval includes zero. Although this does not prove that
the published result is wrong, it does tell us that the answer depends on exactly which comparison and analysis are being used. (Note that their paper also examines rounds, weight classes, and contests between athletes of similar ability.)

Not only does this once again highlight that doing data analysis requires making an argument, but it also raises another question: which analysis was specified because it follows from the argument, and which analyses became important after looking at the data? 

# POSSIBLE ASIDE: Testing the Difference Against Zero

A confidence interval asks whether zero is a plausible value for the population difference. A hypothesis test asks a closely related question:

> If red and blue were equally likely to win, how often would chance produce a difference at least as large as the one we observed?

Under this **null hypothesis**, each bout is like a fair coin flip: red has a 50 percent chance of winning and blue has a 50 percent chance.

```{r observed-difference-object}
ObservedDifference = athens |>
  summarize(Difference = mean(red_win) - mean(blue_win)) |>
  pull(Difference)
```

Simulate 1,000 imaginary Olympics in which color has no advantage.

```{r null-loop}
NullDifferences = NULL
fair_coin = tibble(red_win = c(0, 1))
set.seed(2300)

for (i in 1:1000) {

  one_null = fair_coin |>
    slice_sample(n = nrow(athens), replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(1 - red_win)
    )

  NullDifferences = bind_rows(NullDifferences, one_null)
}
```

```{r inspect-null}
summary(NullDifferences)
```

The null differences are centered near zero because they come from a world in
which red and blue are equally likely to win. We now count how many imaginary
Olympics produced a difference at least as far from zero as the observed
difference, in either direction.

```{r p-value}
NullDifferences |>
  summarize(
    p_value = mean(abs(Difference) >= abs(ObservedDifference))
  )
```

This simulated **p-value** is the fraction of fair-coin results at least as extreme as the observed result. 
A small p-value is evidence against the no-difference hypothesis. 
It is not the probability that the null hypothesis is true.

# What Happens When We Add More Data?

The Athens result comes from one Olympic Games.  But we also have comparable results from Olympics held between 1996 and 2020.

```{r load-all-olympics}
olympics = read_csv(
  "olympics.csv"
) |>
  mutate(
    red_win = if_else(winner == "Red", 1, 0),
    blue_win = if_else(winner == "Blue", 1, 0)
  )
```

Calculate the difference for each Olympic year.

```{r differences-by-year}
olympics |>
  group_by(year) |>
  summarize(
    N = n(),
    Difference = mean(red_win) - mean(blue_win)
  )
```

The result is not stable across Olympics. In some years red wins more; in
others, blue wins more.

Now calculate the difference using all of the contests.

```{r pooled-difference}
olympics |>
  summarize(
    RedWinRate = mean(red_win),
    BlueWinRate = mean(blue_win),
    Difference = RedWinRate - BlueWinRate
  )
```

The apparent red advantage is much smaller when we include more data. Use the
same bootstrap one more time.

```{r all-years-bootstrap}
AllYearsDifferences = NULL
set.seed(42)

for (i in 1:1000) {

  one_difference = olympics |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(blue_win)
    )

  AllYearsDifferences = bind_rows(
    AllYearsDifferences,
    one_difference
  )
}

AllYearsDifferences |>
  summarize(
    Lower = quantile(Difference, .025),
    Estimate = mean(Difference),
    Upper = quantile(Difference, .975)
  )
```

Does this interval include zero? What has changed since we looked only at Athens in 2004?

What if we looked by sport?

```{r}
olympics |>
  group_by(sport) |>
  summarize(
    RedWinRate = mean(red_win),
    BlueWinRate = mean(blue_win),
    Difference = RedWinRate - BlueWinRate
  )
```

If we want to see if these differences were similar by sport, what would change?

```{r}
AllYearsBySportDifferences = NULL
set.seed(42)

for (i in 1:1000) {

  one_difference = olympics |>
    slice_sample(prop = 1, replace = TRUE) |>
    summarize(
      Difference = mean(red_win) - mean(blue_win)
    )

  AllYearsDifferences = bind_rows(
    AllYearsDifferences,
    one_difference
  )
}

AllYearsDifferences |>
  summarize(
    Lower = quantile(Difference, .025),
    Estimate = mean(Difference),
    Upper = quantile(Difference, .975)
  )
```

What does that suggest?

## Empirical Bootstrap

What we have been doing is called the [empirical bootstrap](https://ocw.mit.edu/courses/mathematics/18-05-introduction-to-probability-and-statistics-spring-2014/readings/MIT18_05S14_Reading24.pdf). It's massively useful, because it can be applied for any summary measure of the data: median, percentiles, and measures like regression coefficients. Here is the summary of steps for the empirical bootstrap:

- Decide on the summary measure to be used for the variable (it doesn't have to be the mean)
- Calculate the summary measure on a small subsample (called the bootstrap sample) of the data
- Repeat step 2 many times (how many? Start with 1000, but more is better.) Compile the estimates. 
- Calculate the percentiles of the bootstrap distribution from the previous step. 
- Describe your uncertainty using those percentiles.


# What Have We Learned?

- A margin of error describes how much an estimate might change across samples.
- Bootstrap resampling imitates repeated samples using the data we have.
- The same procedure works for a poll percentage or a difference of means.
- A confidence interval helps us ask whether zero is plausible.
- A hypothesis test asks how surprising our estimate would be if the true
  difference were zero.
- Sampling uncertainty is only one kind of uncertainty; resampling cannot fix
  a sample that poorly represents the population.
- A finding from one dataset can weaken when more evidence becomes available.

# New R Functions Introduced in Lecture 6

Functions introduced in earlier lectures are intentionally omitted.

| Function | What it does |
|---|---|
| `slice_sample()` | Randomly draws rows from a data frame. |
| `set.seed()` | Makes random results reproducible. |
| `tibble()` | Creates a data frame in tidyverse format. |
| `bind_rows()` | Adds rows to the bottom of a data frame. |
| `quantile()` | Returns values at specified percentiles. |
| `abs()` | Returns the absolute distance from zero. |
| `nrow()` | Returns the number of rows in a data frame. |
| `for` | Repeats a block of code a specified number of times. |

# Works Cited

Hill, Russell A., and Robert A. Barton. 2005. “Red Enhances Human Performance
in Contests.” *Nature* 435: 293.
[https://doi.org/10.1038/435293a](https://doi.org/10.1038/435293a).

Peperkoorn, Leonard S., Russell A. Hill, Robert A. Barton, and Thomas V. Pollet.
2024. “Meta-analysis of the Red Advantage in Combat Sports.” *Scientific
Reports* 14: 30822.
[https://doi.org/10.1038/s41598-024-81373-3](https://doi.org/10.1038/s41598-024-81373-3).
Data and code: [https://osf.io/fqkcw/](https://osf.io/fqkcw/).

Schaffner, Brian, Marissa Shih, Stephen Ansolabehere, and Jeremy Pope. 2025.
“Cooperative Election Study Common Content, 2024.” Harvard Dataverse, V9.
[https://doi.org/10.7910/DVN/X11EP6](https://doi.org/10.7910/DVN/X11EP6).
